---
title: "Statistical Inference-Project Part 2"
author: "Praveen Balasubramanian"
date: "Tuesday, March 17, 2015"
output: html_document
---
##PART 2

Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package.

1. Load the ToothGrowth data and perform some basic exploratory data analyses 

```{r}
library(gplots)
library(lattice)
data(ToothGrowth)
nrow(ToothGrowth)
ncol(ToothGrowth)
str(ToothGrowth)
```

```{r}
ToothGrowth$dose #We must consider 0.5, 1 and 2 as scales
ToothGrowth$dose = factor(ToothGrowth$dose, levels=c(0.5,1.0,2.0),
                          labels=c("low","med","high"))
attach(ToothGrowth)
table(supp,dose)
```

```{r, echo=FALSE}
bwplot(len ~ dose | supp)
```



2. Provide a basic summary of the data.

```{r}
summary(ToothGrowth)
aggregate(len,list(supp,dose), mean)
aggregate(len,list(supp,dose), sd)
```

3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose. (Use the techniques from class even if there's other approaches worth considering)

Here, i've chosen to work with a factor of (dose) hence i prefer to go ahead with Anova test as they may provide deeper insight. We then use Tukey's HSD (honest significant difference) test,or the Tukey-Kramer method, which is a single-step multiple comparison procedure and statistical test. It can be used on raw data or in conjunction with an ANOVA (Post-hoc analysis) to find means that are significantly different from each other. 

```{r}                 
anova <- aov(len ~ supp * dose, data=ToothGrowth)
TukeyHSD(anova)
```

```{r, echo=FALSE}
plotmeans(len~interaction(supp,dose,sep=" "),connect=list(c(1,3,5),c(2,4,6)),col=c("red","blue"),xlab="Dose and Supp Combination",main = "Interaction plot with 95% confidence intervals")
```

4. State your conclusions and the assumptions needed for your conclusions. 

Assumption: The sample components are independent and identically distributed.

Conclussion: The Tukey contrast reveals details that there are no statistical evidence for other 'noisy' variables which are relevant, therefore it is evident that changes in dose and size (independent variables) won't affect lenght (dependent variable).
